Using a Highly Dynamic Language for Development

### 2021-10-09 gic2021-highly-dynamic/paper.tex
Paper updates

\begin{figure}[h]
  \centering
  \includegraphics[width=0.4\textwidth]{method combination.png}
  \caption{The standard method combination behaviour illustrated}
  \label{fig:method combination}
\end{figure}

@@

You can also declare the type of a function, to ensure that the compiler can infer return values and arguments across function boundaries. SBCL's block compilation feature, when activated, also allows it to reason across function boundaries within the same compilation unit on its own. We have not yet properly explored block compilation in our code base, as block compilation has a few provisions that we do not fulfil yet.

@@

When focusing on the SBCL implementation, there are several other optimisations that can be done. SBCL offers full access to its code optimisation facilities and even its assembler routines. With enough effort, code can be optimised on the assembly level, and one could even emit special assembly instructions where needed. Some libraries already make use of these systems to speed up computations, but we have not touched them much ourselves. We do however make use of the \code{disassemble} function to check the generated assembly of functions, as another aid in optimisation. We also make use of the integrated statistical profiler to observe the runtime of the game and determine choke points.

@@

The GC we use is SBCL's default GC, a generational, compacting, single-threaded stop-the-world collector. We do hope that SBCL will receive a better GC at some point, but for now we've been focusing on minimising garbage production in our code to avoid frequent and extensive GCs.

@@

\section{Conclusion}
We believe that Common Lisp offers a set of unique features that makes it a great candidate for the kind of highly iterative development processes often required for games.

The expressiveness of the Common Lisp Object System allows for rapid changes in game object behaviour, as well as the creation of reusable pieces of behaviour that can often be seamlessly combined together. With extensions to the Meta Object Protocol, this can even be extended to graphics rendering routines and shaders.

>>>>

\begin{figure}[h]
  \centering
  \includegraphics[width=0.45\textwidth]{method combination.png}
  \caption{The standard method combination behaviour illustrated}
  \label{fig:method combination}
\end{figure}

@@

You can also declare the type of a function, to ensure that the compiler can infer return values and arguments across function boundaries. SBCL's block compilation feature, when activated, also allows it to reason across function boundaries within the same compilation unit on its own.

@@

When focusing on the SBCL implementation, there are several other optimisations that can be done. SBCL offers full access to its code optimisation facilities and even its assembler routines. With enough effort, code can be optimised on the assembly level, and one could even emit special assembly instructions where needed. Some libraries already make use of these systems to speed up computations, but we have not touched them much ourselves due to a lack of time so far. We do however make use of the \code{disassemble} function to check the generated assembly of functions, as another aid in optimisation. We also make use of the integrated statistical profiler to observe the runtime of the game and determine choke points.

@@

The GC we use is SBCL's default GC, a generational, compacting, single-threaded stop-the-world collector. We do hope that SBCL will receive a better GC such as the Memory Pool System\cite{brooksby2002memory} at some point, but for now we've been focusing on minimising garbage production in our code to avoid frequent and extensive GCs.

@@

\section{Conclusion}
The expressiveness of the Common Lisp Object System allows for rapid changes in game object behaviour, as well as the creation of reusable pieces of behaviour that can often be seamlessly combined together. With extensions to the Meta Object Protocol, this can even be extended to graphics rendering routines and shaders.
